On the number of tiles visited by a\\ line segment on a rectangular grid

### 2021-08-24 paper.tex
Change introduction and section structure to make room for the new results on the probability of visiting the maximum number of tiles

Given $a, b \in \mathbb R^+$, consider a grid on $\mathbb R^2$ formed by rectangular \emph{tiles} of width $a$ and height $b$. A line segment of length $\len \in \mathbb R^+$ is located on the plane with arbitrary position and orientation. The segment is said to \emph{visit} a tile if it intersects its interior.\footnote{
The definition uses the interior of the tile, excluding the border, to avoid uninteresting results such as a ``zero-length'' segment visiting (a vertex of) $4$ tiles.} This paper deals with the following problems:
\begin{itemize}
\item What is the maximum number of tiles that the segment can visit?
\item What length should a segment have to visit a given number of tiles?
\item What is the average number of tiles visited by a uniformly random segment of a given length?
\end{itemize}

@@

The problem of how many tiles a \emph{random} segment of a given length visits on average is precisely defined as follows. By symmetry, one endpoint of the segment can be assumed to lie in a fixed, reference tile. The position of this endpoint is \emph{uniformly} distributed on the tile. The segment orientation has a \emph{uniform} distribution on the set of all possible directions, and is \emph{independent} of the endpoint position. As will be seen, solving this problem also answers the inverse question (segment length to visit a given number of tiles on average), which thus need not be considered separately.

@@

The rest of the paper is organized as follows. Fundamental results are presented in \S\ref{part: fund results}, which form the basis of the subsequent analysis. The direct and inverse problems for a deterministic segment are considered in \S\ref{part: max}, first for real-valued lengths, and then for integer-valued lengths on a unit square grid. The problem of computing the average number of tiles visited by a random segment is analyzed in \S\ref{part: ave}.

@@

\section{Average number of tiles visited by a random segment}
\label{part: ave}

@@

Let $\funta: \mathbb R^+ \to \mathbb R^+$ be defined such that $\funta(\len)$ gives the average number of tiles visited by a random segment of length $\len$, with the above distributions.

@@

\begin{proposition}
\label{prop: Pr i}
Given $a, b, \len \in \mathbb R^+$, consider a grid with parameters $a, b$ and a uniformly random segment of length $\len$, as defined above. Let the random variables $i$, $j$ represent the normalized dimensions of the canonical rectangle. For $n \in \mathbb N$,
\begin{equation}
\label{eq: Pr i geq n}
\Pr[i \geq n] = \begin{cases}
\displaystyle
1 &\text{if\ \ } \displaystyle n =1 \\[1 mm]
\displaystyle
\frac{2\len}{\pi a} \left(f\left(\frac{a(n-1)}{\len}\right)-f\left(\frac{a(n-2)}{\len} \right)\right) &\text{if\ \ } \displaystyle 2 \leq n < \frac \len a + 1 \\[4 mm]
\displaystyle
\frac{2\len}{\pi a}\,\, \left(1 - f\left(\frac{a(n-2)}{\len}\right)\right) &\text{if\ \ } \displaystyle \frac\len a + 1 \leq n < \frac\len a+2 \\[3 mm]
\displaystyle
0 &\text{if\ \ } \displaystyle \frac \len a + 2 \leq  n; \\
\end{cases}
\end{equation}
and $\Pr[j \geq n]$ is given by the same expressions with $a$ replaced by $b$.
\end{proposition}

@@

>>>>

Given $a, b \in \mathbb R^+$, consider a grid on $\mathbb R^2$ formed by rectangular \emph{tiles} of width $a$ and height $b$. A line segment of length $\len \in \mathbb R^+$ is located on the plane with arbitrary position and orientation. The segment is said to \emph{visit} a tile if it intersects its interior.\footnote{
The definition uses the interior of the tile, excluding the border, to avoid uninteresting results such as a ``zero-length'' segment visiting (a vertex of) $4$ tiles.} This paper deals with the following problems:
\begin{itemize}
\item What is the maximum number of tiles that the segment can visit?
\item What length should a segment have to visit a given number of tiles?
\item What is the average number of tiles visited by a uniformly random segment of a given length? How often does the random segment visit the maximum number of tiles?
\end{itemize}

@@

The problem of how many tiles a \emph{random} segment of a given length visits is precisely defined as follows. By symmetry, one endpoint of the segment can be assumed to lie in a fixed, reference tile. The position of this endpoint is \emph{uniformly} distributed on the tile. The segment orientation has a \emph{uniform} distribution on the set of all possible directions, and is \emph{independent} of the endpoint position. As will be seen, solving the problem of how many tiles the segment visits on average also answers the inverse question (segment length to visit a given number of tiles on average). In addition, it is interesting to know the probability that the segment visits the maximum number of tiles.

@@

The rest of the paper is organized as follows. Fundamental results are presented in \S\ref{part: fund results}, which form the basis of the subsequent analysis. The direct and inverse problems for a deterministic segment are considered in \S\ref{part: max}, first for real-valued lengths, and then for integer-valued lengths on a unit square grid. The analysis for the random segment is carried out in \S\ref{part: rand}. The average number of tiles is computed in general, and the probability that the segment visits the maximum number of tiles is obtained for a square grid.

@@

\section{Probabilistic characterization for a random segment}
\label{part: rand}

@@

\section{Arbitrary grid with real-valued lengths}

Let $\funta: \mathbb R^+ \to \mathbb R^+$ be defined such that $\funta(\len)$ gives the average number of tiles visited by a random segment of length $\len$, with the distributions specified in \S\ref{part: rand}.

@@

% ***This will go after Theorem~\ref{theo: funta, form}
\begin{proposition}
\label{prop: Pr i}
Given $a, b, \len \in \mathbb R^+$, consider a grid with parameters $a, b$ and a uniformly random segment of length $\len$, as defined above. Let the random variables $i$, $j$ represent the normalized dimensions of the canonical rectangle. For $n \in \mathbb N$,
\begin{equation}
\label{eq: Pr i geq n}
\Pr[i \geq n] = \begin{cases}
\displaystyle
1 &\text{if\ \ } \displaystyle n =1 \\[1 mm]
\displaystyle
\frac{2\len}{\pi a} \left(f\left(\frac{a(n-1)}{\len}\right)-f\left(\frac{a(n-2)}{\len} \right)\right) &\text{if\ \ } \displaystyle 2 \leq n < \frac \len a + 1 \\[4 mm]
\displaystyle
\frac{2\len}{\pi a}\,\, \left(1 - f\left(\frac{a(n-2)}{\len}\right)\right) &\text{if\ \ } \displaystyle \frac\len a + 1 \leq n < \frac\len a+2 \\[3 mm]
\displaystyle
0 &\text{if\ \ } \displaystyle \frac \len a + 2 \leq  n; \\
\end{cases}
\end{equation}
and $\Pr[j \geq n]$ is given by the same expressions with $a$ replaced by $b$.
\end{proposition}

@@

***Clarify here that computing the probability of visiting the maximum number of tiles in the general case is a difficult problem, and it will be dealt with in the square-grid case only

\section{Square grid with real-valued lengths}

***
